---
title: "INFO523 SVM & ANN"
author: "Noah Giebink & Sebastian Deimen"
date: "4/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(kernlab)
library(e1071)
library(tidyverse)
library(knitr)
library(kableExtra)
library(rpart)
library(performanceEstimation)
```

# Classify track explicitness using sociopolitical variables (SVM)

## Data Preprocessing


```{r, message=FALSE}
spot <- read_csv('../data/spot_clean.csv') %>% select(country, track.explicit, happiness,
                                                      median_age, percent_urban,
                                                      percent_internet_users, density_sqkm,
                                                      freedom, gdp)
```

Table 1. There is only one observation of sociopolitical variables per country, but ~50 unique tracks. Therefore, both levels of track.explicit (T/F) are associated with many duplicate observations, making classification difficult. To illustrate, we show the first handful of observations for the Dominican Republic.

```{r, echo=FALSE}
kable(head(spot)) %>%
  kable_styling(latex_options = 'scale_down')

```


```{r}
exp <- group_by(spot, country, track.explicit) %>% tally() %>% filter(track.explicit==TRUE) %>%
  select(country, num_explicit = n)
songs_count <- group_by(spot, country) %>% summarise(count = n()) 
exp_count <- left_join(exp, songs_count, by = 'country')
exp_freq <- mutate(exp_count, freq_explicit = num_explicit/count) %>% select(country, freq_explicit)
exp_freq <- mutate(exp_freq, explicit_label = ifelse(freq_explicit > median(exp_freq$freq_explicit), 1, -1))

spot_exp <- left_join(exp_freq, spot, by = 'country') %>% select(-track.explicit, -freq_explicit)
spot_exp <- unique(spot_exp)
spot_exp$'explicit_label' = factor(spot_exp$explicit_label)

```

Table 2. We solve this problem by labelling countries' taste as particularly explicit (+1) or not particularly explicit (-1) and retaining only unique observations (i.e. only one observation per country).

```{r, echo=FALSE}
kable(spot_exp)%>%
  kable_styling(latex_options = 'scale_down')

```


```{r}
spot_exp <- ungroup(spot_exp)
perf <- performanceEstimation(PredTask(explicit_label~happiness+median_age+percent_urban+
                                         percent_internet_users+density_sqkm+freedom+gdp, 
                               data = spot_exp),
                               Workflow(learner = 'svm'), 
                               EstimationTask(metrics = 'acc',
                               method=CV(nReps= 5, nFolds = 5)))
summary(perf)

# add best.tune to get best performing model

# ADD train/test split or cross-validation to get model performance

# build SVMs
svm_data <- select(spot_exp, -country)
exp_svm <- svm(explicit_label~., data = svm_data, cost = 0.01, gamma = 0.001)
exp_svm2 <- svm(explicit_label~., data = svm_data, cost = 100, gamma = 100)
summary(exp_svm)
summary(exp_svm2)

# predict

# model performance

```
